---
classoption: b5paper
geometry: margin=1in
header-includes:
- \usepackage{multirow}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage[table]{xcolor}
- \usepackage{tabularx,booktabs}
- \usepackage{boxedminipage}
- \usepackage{graphicx}
- \usepackage{rotating}
- \usepackage{longtable}
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{tabular}{l!{\color{green}\vrule}l}
    \arrayrulecolor{red}\hline
    test & test\\\arrayrulecolor{blue}\hline
\end{tabular}

\begin{table}[!hbtp]
\begin{center}
\begin{normalsize}
\color{black}
\begin{tabular}{!{\color{orange}\vrule}r!{\color{orange}\vrule}r!{\color{orange}\vrule}r!{\color{orange}\vrule}r!{\color{orange}\vrule}r!{\color{orange}\vrule}l!{\color{orange}\vrule}}
\arrayrulecolor{orange}
\hline
&Sepal.Length&Sepal.Width&Petal.Length&Petal.Width&Species\\ 
\hline
1&5.10&3.50&1.40&0.20&setosa\\ 
\hline
2&4.90&3.00&1.40&0.20&setosa\\ 
\hline
3&4.70&3.20&1.30&0.20&setosa\\ 
\hline
4&4.60&3.10&1.50&0.20&setosa\\ 
\hline
5&5.00&3.60&1.40&0.20&setosa\\ 
\hline
\end{tabular}
\end{normalsize}
\end{center}
\end{table}
\color{black}

\definecolor{platinum}{rgb}{0.9, 0.89, 0.89} 
\begin{table}[!hbtp]
\begin{center}
\begin{normalsize}
\color{black}
\begin{tabular}{!{\color{orange}\vrule}r!{\color{orange}\vrule}r!{\color{orange}\vrule}r!{\color{orange}\vrule}r!{\color{orange}\vrule}r!{\color{orange}\vrule}l!{\color{orange}\vrule}}
\arrayrulecolor{orange}
\hline
\cellcolor{platinum}&\cellcolor{platinum}\textbf{Sepal.Length}&\cellcolor{platinum}\textbf{Sepal.Width}&\cellcolor{platinum}\textbf{Petal.Length}&\cellcolor{platinum}\textbf{Petal.Width}&\cellcolor{platinum}\textbf{Species}\\
\hline
\rowcolor{red}1&5.10&3.50&1.40&0.20&setosa\\ 
\hline
\cellcolor{platinum}2&\cellcolor{platinum}4.90&\cellcolor{platinum}3.00&\cellcolor{platinum}1.40&\cellcolor{platinum}0.20&\cellcolor{platinum}setosa\\ 
\hline
3&4.70&3.20&1.30&0.20&setosa\\ 
\hline
\rowcolor{platinum}4&4.60&3.10&1.50&0.20&\color{red}setosa\\ 
\hline
5&5.00&3.60&1.40&0.20&setosa\\ 
\hline
\end{tabular}
\end{normalsize}
\end{center}
\end{table}
\color{black}

```{r,results='asis'}
require(ztable)
z=ztable(head(iris),zebra=1)
cgroup=c("Sepal","Petal","Species")
n.cgroup=c(2,2,1)
z=addcgroup(z,cgroup=cgroup,n.cgroup=n.cgroup,color="red")
# cgroup2=c("Length","Species")
# n.cgroup2=c(4,1)
# z=addcgroup(z,cgroup=cgroup2,n.cgroup=n.cgroup2,top=TRUE)
rgroup=c("OneToThree","Four","FiveToSix")
n.rgroup=c(3,1,2)
z=addrgroup(z,rgroup=rgroup,n.rgroup=n.rgroup,color=c("blue","red","yellow"))
print(z,type="latex")

```


