---
title: "Make a Heatmap Table using `ztable`"
author: "Keon-Woong Moon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{heatmapTable}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA,
  message=FALSE
)
```

A heat map (or heatmap) is a graphical representation of data where the individual values contained in a matrix are represented as colors. You can summarize the the diagnosis and smoking status of 857 patients of acute coronary syndrome(acs) using table() function.  

```{r}
require(moonBook)
x=table(acs$Dx,acs$smoking)
x
```

## Basic Table

You can make `html` or `LaTex` table easly with ztable.
```{r,results="asis"}
library(ztable)
library(magrittr)
options(ztable.type="html")
z=ztable(x) 
print(z,caption="Table 1. Basic Table")
```

## Formatting the Table

You can change the background color and font color of `ztable` using addCellColor function. For example, you can change the cell color of the 3rd row, 2nd column. Please keep in mind that the ztable count colname and rowname. 

```{r,results="asis"}
z %>% 
    addCellColor(4,3,bg="orange",color="white") %>% 
    print(caption="Table 2. Add Cell Color")
```


## Conditional Formatting

You can select rows with logical expression. You can select cols with column name.

```{r,results='asis'}
ztable(head(iris),caption="Table 3. Conditinoal Formatting: Sepal.Width >= 3.5") %>%
    addRowColor(rows=1,bg="#C90000",color="white") %>%
    addCellColor(condition=Sepal.Width>=3.5,cols=Sepal.Width,color="red") 
```

##  Make a Heatmap Table

You can make a heatmap table in which background colors representing the values. With makeHeatmap() function, you can make a heatmap table easily. The namakeHeatmap() function apply the "Reds" palette from RColorBrewer package.

```{r,results="asis"}
z %>% makeHeatmap() %>% print(caption="Table 4. Heatmap Table")
```

## Heatmap Table with desired palette

You can change the palette with palette argument. For example, you can use the "Blue" palette.

```{r,results='asis'}
ztable(head(mtcars)) %>% 
    makeHeatmap(palette="Blues") %>%
    print(caption="Table 5. Heatmap table with 'Blue' palette")
```

## Heatmap Table with user-defined palette

With the gradientColor() function, you makes sequential colour graient palette easily. 

```{r,fig.width=8,out.width='100%'}
mycolor=gradientColor(low="yellow",mid="orange",high="red",n=20,plot=TRUE)
mycolor
```

```{r,results='asis'}
ztable(head(mtcars[1:5])) %>% 
    makeHeatmap(mycolor=mycolor) %>%
    print(caption="Table 6. Heatmap table with user-defined palette")
```

## Heatmap Table with non-numeric data

You can make heatmap table with data containing non-numeric columns. Only columns with numeric data affected by this function.

```{r,results='asis'}
ztable(head(acs[1:10])) %>% 
    makeHeatmap %>%
    print(caption="Table 7. Heatmap table with non-numeric data")
```
